---
title: "Model Comparison Table 1: Kriging + Cosine Mixture Function"
author: "Ellen Wei"
date: "2024-06-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Imports
```{r}
library(mlrMBO)
library(lhs)
library(smoof)
library(ggplot2)
library(DiceKriging)
library(gridExtra)
library(GGally)
```

1. Define objective function and its parameters using the package smoof.
  Cosine mixture function (1D), Branin (2D), Hartmann(6D)
  
2. Generate initial design (optional).
  LHS, Random, Uniform
  
3. Define mlr learner for surrogate model (optional).
  Kriging (default), Random Forest
  
4. Set up a MBO control object.

5. Start the optimization with run=mbo().

6. Visualize runs: exampleRun, plotExampleRun

## General function and parameters

```{r}

# Cosine Mixture function
fun.cosine = makeCosineMixtureFunction(1)
fun.cosine = convertToMinimization(fun.cosine)
#print(autoplot(fun.cosine, show.optimum = TRUE))
print(fun.cosine)

# Number of run iterations
multi_iters = 15
control_iters = 15

  ### !!! NOTE: control_iters > multi_iters or things will not plot !!!

# Random LHS initial design parameters
des.n = 5

  # Random LHS Cosine mixture function generate
des.cosine_random = generateDesign(des.n, par.set = getParamSet(fun.cosine), randomLHS)

  # Maximin LHS Cosine mixture 
des.cosine_maximin = generateDesign(des.n, par.set = getParamSet(fun.cosine), maximinLHS)

  # Optimum LHS Cosine mixture
des.cosine_optimum = generateDesign(des.n, par.set = getParamSet(fun.cosine), optimumLHS)

# Kriging model (surrogate) constructor

surr.km = makeLearner("regr.km", predict.type = "se", covtype = "matern3_2", control = list(trace = FALSE))

```


```{r}
# separately generate obj.fun, surr.km (use makeLearner), des (LHS, random, uniform)


generate_model <- function(obj.fun, des, des_name, surr.km, control.iters=control.iters, control.crit){
  # Generates model but doesn't run or visualize
  # obj.fun: objective function ex. makeBraninFunction()
  # des required variables:
    # n: number of initial points
    # fun: initial design function ex. lhs::randomLHS
  # surr.km: surrogate model ex. makeLearner("regr.km", predict.type = "se", covtype = "matern3_2", control = list(trace = FALSE))
  
  # control: infill criteria
    # control.iters
    # control.crit: infill criteria ex. makeMBOInfillCritEI()
  
  des$y = apply(des, 1, obj.fun)
  control = makeMBOControl()
  control = setMBOControlTermination(control, iters = control.iters)
  control = setMBOControlInfill(control, crit = control.crit)
  
  control.crit_name = switch(as.character(substitute(control.crit)), 
                             "makeMBOInfillCritEI"="EI",
                             "makeMBOInfillCritCB"="CB",
                             "makeMBOInfillCritAEI"="AEI",
                             "makeMBOInfillCritStandardError"="SE")
  
  model_name = paste(control.crit_name, "+", des_name)
  
  return(list(obj.fun=obj.fun, learner=surr.km, des=des, control=control,
              model_name=model_name))
}

run_model <- function(model, show.info=FALSE, multi_iters= multi_iters) {
  # Runs model for _ iterations and displays plots
  
  run = exampleRun(fun=model$obj.fun, design=model$des, learner=model$learner, control=model$control,
                   show.info=show.info)
  plotExampleRun(run, iters=multi_iters)
  
  return(run)
} 

get_results <- function(run){
  return(as.data.frame(run$mbo.res$opt.path))
}

get_best <- function(run){
  return(run$mbo.res$best.ind)
}

get_name <- function(model){
  return(model$model_name)
}
```

# Table 1: Kriging+Cosine Mixture Function (1D)


## 1. EI, Random

```{r}

model_1 = generate_model(fun.cosine, des.cosine_random, "Random", surr.km, 
                         control.iters=control_iters, control.crit=makeMBOInfillCritEI())

# optional: define multi_iters here. default is c(1L, 2L, 10L)

run_1 = run_model(model=model_1, multi_iters=multi_iters)

results_1 = get_results(run_1)
results_1

```

## 2. CB, Random

```{r}
set.seed(123)
model_2 = generate_model(fun.cosine, des.cosine_random, "Random", surr.km,
                         control.iters=15, control.crit=makeMBOInfillCritCB())

run_2 = run_model(model=model_2, multi_iters=15)

results_2 = get_results(run_2)
results_2
```


## 3. AEI, Random

```{r}
model_3 = generate_model(fun.cosine, des.cosine_random, "Random", surr.km,
                         control.iters=control_iters, control.crit=makeMBOInfillCritAEI())

run_3 = run_model(model=model_3, multi_iters=multi_iters)

results_3 = get_results(run_3)
results_3

```

## 4. EI, Maximin

```{r}

model_4 = generate_model(fun.cosine, des.cosine_maximin, "Maximin", surr.km, 
                         control.iters=control_iters, control.crit=makeMBOInfillCritEI())

run_4 = run_model(model=model_4, multi_iters=multi_iters)

results_4 = get_results(run_4)
results_4

```

## 5. CB, Maximin

```{r}

model_5 = generate_model(fun.cosine, des.cosine_maximin, "Maximin", surr.km,
                         control.iters=control_iters, control.crit=makeMBOInfillCritCB())

run_5 = run_model(model=model_5, multi_iters=multi_iters)

results_5 = get_results(run_5)
results_5

```


## 6. AEI, Maximin

```{r}
model_6 = generate_model(fun.cosine, des.cosine_maximin, "Maximin", surr.km,
                         control.iters=control_iters, control.crit=makeMBOInfillCritAEI())

run_6 = run_model(model=model_6, multi_iters=multi_iters)

results_6 = get_results(run_6)
results_6

```

## 7. EI, Optimum

```{r}
model_7 = generate_model(fun.cosine, des.cosine_optimum, "Optimum", surr.km, 
                         control.iters=control_iters, control.crit=makeMBOInfillCritEI())

run_7 = run_model(model=model_7, multi_iters=multi_iters)

results_7 = get_results(run_7)
results_7

get_best(run_7)

```


## 8. CB, Optimum

```{r}
model_8 = generate_model(fun.cosine, des.cosine_optimum, "Optimum", surr.km,
                         control.iters=control_iters, control.crit=makeMBOInfillCritCB())

run_8 = run_model(model=model_8, multi_iters=multi_iters)

results_8 = get_results(run_8)
results_8

get_best(run_8)
```


## 9. AEI, Optimum

```{r}
model_9 = generate_model(fun.cosine, des.cosine_optimum, "Optimum", surr.km,
                         control.iters=control_iters, control.crit=makeMBOInfillCritAEI())

run_9 = run_model(model=model_9, multi_iters=multi_iters)

results_9 = get_results(run_9)
results_9

get_best(run_9)
```

## Table 1 Results

```{r}
# rows: model
# columns: best.ind, train.time, propose.time, se

# evaluation functions
get_train.time <- function(results){
  return(sum(na.omit(results$train.time)))
}

get_propose.time <- function(results){
  return(sum(na.omit(results$propose.time)))
}

get_se <- function(results){
  return(sum(na.omit(results$se)**2))
}

make_table_row <- function(model, run, results) {
  name <- get_name(model)
  best <- round(get_best(run), 3)
  train_time <- round(get_train.time(results), 3)
  propose_time <- round(get_propose.time(results), 3)
  se <- round(get_se(results), 3)
  
  return(c(name, best, train_time, propose_time, se))
}


models <- list(model_1, model_2, model_3, model_4, model_5, model_6, model_7, model_8, model_9)
runs <- list(run_1, run_2, run_3, run_4, run_5, run_6, run_7, run_8, run_9)
results <- list(results_1, results_2, results_3, results_4, results_5, results_6, results_7, results_8, results_9)

# Apply make_table_row to each element of models, runs, and results
table_rows <- mapply(make_table_row, models, runs, results, SIMPLIFY = FALSE)

# Combine the table rows into a data frame
table_1 <- as.data.frame(do.call(rbind, table_rows))
colnames(table_1) <- c("Model", "Best Iteration", "Train Time", "Propose Time", "Sum SE")

table_1
```


